---
title: "Homework 1"
author: "Jonah Schrager"
date: "7/8/2021"
output: word_document
---

#Question 2
```{r}
setwd("~/Desktop/Data Mining/Lecture 1")
data <- read.csv("HW1data.csv")
  str(data) #looks at structure of data (Lecture 1)
  library(tidyverse)  
```
#Question 2A- ggplot- Best fitting line with OLS
```{r}
ggplot(data, aes(x = Ht, y = LBM)) + 
  geom_point() +
geom_smooth(method = "lm", se = FALSE) #adding best fitting line
```
#question 2B # For every .66 increase in LBM, there is an increase in 1 cm in Ht

```{r}
# Multiple Linear Regression for LBM and Ht
fit <- lm(Ht ~ LBM, data = data)
summary(fit)  #show results for LBM and Ht
```
#Question 2C # # For every 2.5 decrease in RCC, there is an increase in 1 cm in Ht
```{r}
# Multiple Linear Regression for LBM and Ht and RCC
fit <- lm(Ht ~ LBM + RCC, data = data)
summary(fit) # show results for LBM and Ht and RCC
```
#Question 2D # for every 1 cm decrease in Ht, there is a increase in male obs by .87
```{r}
# Multiple Linear Regression for LBM and Ht and RCC and Sex
fit <- lm(Ht ~ LBM + RCC + Sex, data = data)
summary(fit) # show results for LBM and Ht and RCC and Sex
```
#Question 2E #There looks like there is a correlation that the males are taller.
```{r}
ggplot(data, aes(x = Ht, y = LBM, color = Sex) ) +
     geom_point() +
     geom_smooth(method = "lm", se = FALSE)
```
#Question 2F
```{r}
obs1 <- head(predict(fit), Sex = f, LBM = 180, RCC = 4.5, n = 1)
```
#Qustion 2G
```{r}
obs2 <- head(predict(fit), Sex = f, LBM = 181, RCC = 4.5, n = 1)

obs2 - obs1

```
#Question 2H i j
```{r}
library(caret)
fitControl <- trainControl(method = "cv", number = 10)
set.seed(1)
summarize(group_by(data, Sex), 
          mean = mean(LBM))
       
set.seed(1)
model.knn.sc <- train(LBM ~ ., 
                  data = data, 
                  method = "knn",
                  preProc = c("center", "scale"),
                  trControl = fitControl) 


model.knn.sc

kGrid <- expand.grid(k = seq(1, 19, by = 2))
set.seed(1)
model.knn.sc.k <- train(LBM ~ ., 
                      data = data, 
                      method = "knn",
                      preProc = c("center", "scale"),
                      trControl = fitControl,
                      tuneGrid = kGrid)

```


#Question 3A
```{r}
set.seed(1)
x1 <- runif(10, 0, 1)
x2 <- runif(10, 0, 1)
y <- 2*x1 + 3*x2 + rnorm(10, 0, 1)
```
#Question 3B
```{r}
sqrt((x1 - .5)^2 + (x2 - .5)^2)
```
#Question 3C
```{r}
x <- mean(y[c(3, 9)])
```
#Question 3D
```{r}
lm.fit <- lm(y ~ x1 + x2)
predict(lm.fit, newdata = data.frame(x1 = .5, x2 = .5))
```
#Question 3E
```{r}
#Seed 2
set.seed(2) 
x1 <- runif(10, 0, 1)
x2 <- runif(10, 0, 1)
y <- 2*x1 + 3*x2 + rnorm(10, 0, 1)

sqrt((x1 - .5)^2 + (x2 - .5)^2)

t <- mean(y[c(1, 3)])

lm.fit <- lm(y ~ x1 + x2)
predict(lm.fit, newdata = data.frame(x1 = .5, x2 = .5))

#Seed 3
set.seed(3) 
x1 <- runif(10, 0, 1)
x2 <- runif(10, 0, 1)
y <- 2*x1 + 3*x2 + rnorm(10, 0, 1)

sqrt((x1 - .5)^2 + (x2 - .5)^2)

z <- mean(y[c(3, 9)])

lm.fit <- lm(y ~ x1 + x2)
predict(lm.fit, newdata = data.frame(x1 = .5, x2 = .5))

var(c(x,y,z))

```
















