---
title: "Assignment 1.0"
author: "Jonah Schrager"
date: "2/11/2022"
output: html_document
---

<style type="text/css">

div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0;
}
h1.title {
  font-size: 24px;
  color: DarkRed;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 { /* Header 1 - and the author and data headers use this too  */
    font-size: 22px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: center;
}
h2 { /* Header 2 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h3 { /* Header 3 - and the author and data headers use this too  */
    font-size: 15px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h4 { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}
</style>

```{r setup, include=FALSE}
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("cowplot")) {
   install.packages("cowplot")
   library(cowplot)
}
if (!require("latex2exp")) {
   install.packages("latex2exp")
   library(latex2exp)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("gapminder")) {
   install.packages("gapminder")
   library(gapminder)
}
if (!require("png")) {
    install.packages("png")             # Install png package
    library("png")
}
if (!require("RCurl")) {
    install.packages("RCurl")             # Install RCurl package
    library("RCurl")
}
if (!require("colourpicker")) {
    install.packages("colourpicker")              
    library("colourpicker")
}
if (!require("gganimate")) {
    install.packages("gganimate")              
    library("gganimate")
}
if (!require("gifski")) {
    install.packages("gifski")              
    library("gifski")
}
if (!require("magick")) {
    install.packages("magick")              
    library("magick")
}
if (!require("grDevices")) {
    install.packages("grDevices")              
    library("grDevices")
}
if (!require("jpeg")) {
    install.packages("jpeg")              
    library("jpeg")
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}
if (!require("patchwork")) {
   install.packages("patchwork")
   library(patchwork)
}
# knitr::opts_knit$set(root.dir = "C:/Users/75CPENG/OneDrive - West Chester University of PA/Documents")
# knitr::opts_knit$set(root.dir = "C:\\STA490\\w05")
##
knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE)
```


# Plot using GGPLOT
```{r}
library(RCurl)
x <- getURL("https://raw.githubusercontent.com/JSchrager94/STA-553/main/Data/penguins.csv?token=GHSAT0AAAAAABQ7IHI3L2FWUPSS334LFIQOYP3B62A")
p <- read.csv(text = x)


bill_flipper_plot <- ggplot(data= p) + geom_point(mapping = aes (bill_length_mm, flipper_length_mm, color= species, shape=species, size=body_mass_g))


penguin_plot <- bill_flipper_plot + labs(x = "Bill Length", y = "Flipper Length", title = "Bill Length vs Flipper Length")
penguin_plot<- penguin_plot + theme_dark() + labs(title = "Bill Length vs Flipper Length", subtitle = "Plot points proportional to body mass",caption = "Jonah Schrager STA 553")




myurl <- "https://i.dailymail.co.uk/i/pix/2015/03/06/2660A3DA00000578-0-The_local_rockhopper_penguins_are_hugely_popular_with_visitors-m-79_1425649402734.jpg"
z <- tempfile()
download.file(myurl,z,mode = "wb")
pic <- readJPEG(z)
raster.penguin <- as.raster(pic)



pic2 <- penguin_plot + 
  inset_element(p = raster.penguin,
                left = .05,
                bottom = .65,
                right = .25,
                top = .95)
pic2
```
