title: ""
author: "Jonah Schrager"
date: "2/28/2022"
output: html_document
---

```{r setup, include=FALSE}
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("cowplot")) {
   install.packages("cowplot")
   library(cowplot)
}
if (!require("latex2exp")) {
   install.packages("latex2exp")
   library(latex2exp)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("gapminder")) {
   install.packages("gapminder")
   library(gapminder)
}
if (!require("png")) {
    install.packages("png")             # Install png package
    library("png")
}
if (!require("RCurl")) {
    install.packages("RCurl")             # Install RCurl package
    library("RCurl")
}
if (!require("colourpicker")) {
    install.packages("colourpicker")              
    library("colourpicker")
}
if (!require("gganimate")) {
    install.packages("gganimate")              
    library("gganimate")
}
if (!require("gifski")) {
    install.packages("gifski")              
    library("gifski")
}
if (!require("magick")) {
    install.packages("magick")              
    library("magick")
}
if (!require("grDevices")) {
    install.packages("grDevices")              
    library("grDevices")
}
if (!require("jpeg")) {
    install.packages("jpeg")              
    library("jpeg")
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}
if (!require("patchwork")) {
   install.packages("patchwork")
   library(patchwork)
}
if (!require("survey")) {
   install.packages("survey)")
   library(survey)
}
knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE)
```


#The four subsets contain information that potentially impacts the life expectancy of the world. The data cover relative information between between 1800 and 2018. 

```{r}
library(RCurl)
x <- getURL("https://raw.githubusercontent.com/JSchrager94/STA-553/main/Data/income_per_person.csv")
income.person <- read.csv(text = x)

x <- getURL("https://raw.githubusercontent.com/JSchrager94/STA-553/main/Data/life_expectancy_years.csv")
life.expectancy <- read.csv(text = x)

x <- getURL("https://raw.githubusercontent.com/JSchrager94/STA-553/main/Data/countries_total.csv")
country.totals <- read.csv(text = x)

x <- getURL("https://raw.githubusercontent.com/JSchrager94/STA-553/main/Data/population_total.csv")
pop.size <- read.csv(text = x)

x <- getURL("https://raw.githubusercontent.com/JSchrager94/STA-553/main/Data/POC.csv")
poc <- read.csv(text = x)
```

#creating longitudinal data for life expectancy
```{r}
life.expectancy.long <- life.expectancy %>%
  gather(key = "Year",       # the column names of the wide table
         value = "lifeExp",  # the numerical values of the table
         -geo,          # drop country variable: its value will not be gathered (stacked)!
         na.rm = TRUE)       # removing records with missing values
##

```

```{r}
correct.life.exp.data <- life.expectancy.long %>%
                      mutate(year = substr(Year,2,5)) %>%
                      select(-Year)

```

#creating longitudinal data for income
```{r}
income.person.long <- income.person %>%
  gather(key = "Year",      
         value = "income", 
         -geo,         
         na.rm = TRUE)      


```

```{r}
correct.income.person.data <- income.person.long %>%
                      mutate(year = substr(Year,2,5)) %>%
                      select(-Year)

```

#creating longitudinal data for population size
```{r}
pop.size.long <- pop.size %>%
  gather(key = "Year",      
         value = "pop_size", 
         -geo,         
         na.rm = TRUE)      


```

#getting rid of x in the year variable
```{r}
correct.pop.size.long <- pop.size.long %>%
                      mutate(year = substr(Year,2,5)) %>%
                      select(-Year)

```


#Sorting data by year
```{r}
#correct.income.person.data[order(correct.income.person.data$year),]

#correct.life.exp.data[order(correct.life.exp.data$year),]
```


# Merge data stes together
```{r}
LifeExpIncom = merge(x = correct.income.person.data, y = correct.life.exp.data)


LifeExpIncom.1 = merge(x = LifeExpIncom, y = country.totals, by.x = 'geo', by.y ='name', all = FALSE)

LifeExpIncom.2 = merge(x = LifeExpIncom.1, y = correct.pop.size.long)
life.exp <- round(LifeExpIncom.2$lifeExp)


```

#subsetting datat set for year 2015
```{r}
year.15 <- subset(LifeExpIncom.2, year == 2015)

```


#interactive plot for 2015
```{r}

pop.size = year.15$`pop size`


plot_ly(
    data = year.15,
    x = ~lifeExp,  # Horizontal axis 
    y = ~income,  
    p = ~pop.size, # Vertical axis 
     color = ~factor(geo), # must be a numeric factor
     text = ~geo,     # show the species in the hover text
     ## using the following hovertemplate() to add the information of the
     ## two numerical variable to the hover text.
     hovertemplate = paste('<i><b>Population<b></i>: %{p}',
                           '<br><b>%{text}</b>'),
     alpha  = 0.5,
     size = ~pop_size,
     type = "scatter",
     mode = "markers"
 ) %>%
    layout(  
      ## graphic size
      with = 700,
      height = 700)
```


```{r}
year <- as.numeric(LifeExpIncom.2$year)

```


#interactive plot for all the years
```{r}
lifeExp_num <- as.data.frame(apply(LifeExpIncom.2, 2, as.numeric))  # Convert all variable types to numeric
sapply(lifeExp_num, class)
lifeExp_num$geo <- as.character(lifeExp_num$geo) 
sapply(lifeExp_num, class)


w <- ggplot(lifeExp_num, aes(income, lifeExp, 
                  size = pop_size, colour = geo)) +
           geom_point(alpha = 0.5, show.legend = FALSE) +
           scale_colour_manual(values = country_colors) +
           #scale_color_manual(values=c("dodgerblue4", "darkolivegreen4","darkorchid3")) +
           #scale_color_brewer(palette="Set1") +
           scale_size(range = c(2,50)) +
           scale_x_log10() +
           # break down the previous single plot by continent 
           # facet_wrap(~continent) +     # create multiple panels according to the continents
           # Here comes the gganimate specific bits
           labs(title = 'Year: {frame_time}', 
                    x = 'Life Expectancy',
                    y = 'Income') +
           transition_time(year) +
           ease_aes('linear')
###
animate(w, renderer = gifski_renderer(),
          rewind = TRUE)

```



#Random sample of 500 gas stations 
```{r}
samp <- poc %>% ungroup() %>% sample_n(500)
```


#gas stations
```{r}
## preparing data
#poc <- read_csv("https://raw.githubusercontent.com/JSchrager94/STA-553/main/Data/POC.csv")
poc.site <- samp[samp$POC == 1,]
# geo styling
geostyle <- list(scope = 'usa',
                 projection = list(type = 'albers usa'),
                 showland = TRUE,
                 landcolor = toRGB("gray95"),
                 subunitcolor = toRGB("gray85"),
                 countrycolor = toRGB("gray85"),
                 countrywidth = 0.5,
                 subunitwidth = 0.5
               )
## plotting map
fig <- plot_geo(samp, lat = ~ycoord, lon = ~xcoord) %>%
       add_markers(text = ~SITE_DESCRIPTION, 
                   color = "red", 
                   symbol = I("circle"), 
                   size = I(8), 
                   hoverinfo = "text" )   %>%
        layout( title = 'Gas stations', geo = geostyle)
fig
```
